---
title: "template_microshades_taxabarplot"
author: "Madi"
date: "2023-05-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(qiime2R)
library(tidyverse)
library(cowplot)
library(magrittr)
library(viridis)
library(microshades)
library(phyloseq)
```

**These Functions Are Not Needed Anymore**
Keeping these for the future in case I need them but I most likely won't. 
```{r}
# ## otu table prep with missing samples 
# ## assumes that you're reading in a 'tsv' file 
# otu_phyloseq_prep <- function(otu_fp, missing_samples){
#   otu_table <- read_tsv(file = otu_fp)
#   otu_table %>% 
#     gather(-asv, key = sample, value = count) %>% 
#     filter(sample %in% missing_sample) %>% 
#     spread(sample, count) -> otu_missing
#   names(otu_missing)[names(otu_missing) == 'asv'] <- ' '
#   otu_missing <- as.matrix(otu_missing)
#   class(otu_missing) <- "numeric"
#   return(otu_missing)
# }
# 
# ## tax table prep with missing samples 
# ## assumes that its a 'qza' file from qiime2
# tax_table_prep <- function(taxonomy_fp, wanted_asvs){
#   tax_table <- read_qza(taxonomy_fp)$data 
#   tax_table %>% 
#     parse_taxonomy() %>% 
#     as_tibble(rownames = 'asv') %>%
#     filter(asv %in% wanted_asvs) -> tax_missing
#   names(tax_missing)[names(tax_missing) == 'asv'] <- ' '
#   return(tax_missing)
# }
# 
# ## putting entire taxa barplot together
# whole_tax_barplot <- function(tax_barplot, tax_barplot_legend){
#   tax_all <- plot_grid(tax_barplot, tax_barplot_legend,  rel_widths = c(1, .25),
#                        align = 'tblr',
#                        axis = 'tblr')
#   return(tax_all)
# }
```

**Overall Functions Needed for the Giant End Function**
These are in order of their use in the end-product function and will be used in the overall snakemake workflow. 
```{r}
## this assumes that you're reading in a qiime2 'qza' file
## this also doesn't filter out anything for you 
## but you usually shouldn't need the filtration steps

micro_file_prep <- function(metadata_fp,
                            otu_table_fp,
                            tax_table_fp){
  ## metadata file prep 
  tmpMeta <- read_tsv(metadata_fp, n_max = 2)
  mycols <- colnames(tmpMeta)
  metadata <- read_tsv(metadata_fp, skip = 2, col_names = mycols)
  names(metadata)[names(metadata) == '#SampleID'] <- 'sampleid'
  metadata <- as.data.frame(metadata)
  rownames(metadata) <- metadata$sampleid
  ## otu file prep
  otu_table <- read_qza(file = otu_fp)$data
  ## taxonomy file prep
  tax_table <- read_qza(file = tax_fp)$data %>% 
    parse_taxonomy()
  tax_table %>% 
    filter(rownames(tax_table) %in% rownames(otu_table)) -> tax_table
  ## creating my named list 
  my_list <- list(Metadata = metadata,
                  OTUTable = otu_table,
                  Taxonomy = tax_table)
  return(my_list)
}
# 
# 
# pseq <- phyloseq(otu_table(pseq_otu, taxa_are_rows = TRUE),
#                  tax_table(as.matrix(pseq_tax)),
#                  sample_data(pseq_meta))

## generating color palettes based on the phyloseq object you generated
## generate pseq like normal since there's already a nice function for it 
## just returns the color_objs_GP for you, you'll need to extract the mdf/cdf for the next function 
mdf_cdf_gp <- function(pseq, tax_table, tax_rank, tax_subgroup){
  pseq %>% 
  tax_glom(tax_subgroup) %>% 
    phyloseq::transform_sample_counts(function(x) {
     x/sum(x)
      }) %>% 
    psmelt() %>% 
  filter(Abundance > 0) -> mdf_prep
  phylum_table <- tax_glom(pseq, taxrank=tax_rank, ) %>% otu_table()
  phyla.otunames <- rownames(phylum_table)
  phylums <- tax_table(as.matrix(tax_table))[phyla.otunames,tax_rank]
  sorted_phylums <- phylums[order(rowSums(phylum_table), decreasing=T)]
## can take ranked phylums out if you don't have issues with top abund. phylums being repeated
## ranked_phylums <- sort(unique(sorted_phylums[40:1]))
  color_objs_GP <- create_color_dfs(mdf_prep,
                                 subgroup_level = tax_subgroup,
                                 selected_groups = sorted_phylums[5:1])
  return(color_objs_GP)
}

tax_barplot <- function(mdf_GP, cdf_GP){
  barplot <- plot_microshades(mdf_GP, cdf_GP)
  barplot_1 <- barplot + scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
    theme_bw(base_size = 18) +
    theme(legend.position = "none") +
    theme(axis.text.x = element_blank())
  return(barplot_1)
}

tax_barplot_legend <- function(mdf_GP, cdf_GP, tax_subgroup){
  GP_legend <-custom_legend(mdf_GP, cdf_GP, 
                          legend_key_size=unit(0.4, "cm"),
                          legend_text_size=15, subgroup_level=tax_subgroup)
  return(GP_legend)
}

```

**Putting all microshades functions into a giant function to create a mini-pipeline**
```{r}
microshades_tax_barplot <- function(meta_fp, 
                                    otu_fp, 
                                    tax_fp,
                                    tax_rank, 
                                    tax_subgroup,
                                    title_text, 
                                    y_text){
  pseq_files <- micro_file_prep(meta_fp,
                                otu_fp,
                                tax_fp)
  pseq_meta <- pseq_files$Metadata
  pseq_otu <- pseq_files$OTUTable
  pseq_tax <- pseq_files$Taxonomy 
  pseq <- phyloseq(otu_table(pseq_otu, taxa_are_rows = TRUE),
                 tax_table(as.matrix(pseq_tax)),
                 sample_data(pseq_meta))
  mdf_and_cdf <- mdf_cdf_gp(pseq, 
                            pseq_tax, 
                            tax_rank, 
                            tax_subgroup)
  mdf_GP <- mdf_and_cdf$mdf
  cdf_GP <- mdf_and_cdf$cdf
  main_barplot <- tax_barplot(mdf_GP, cdf_GP) +
    ggtitle(title_text) +
    ylab(y_text)
  legend <- tax_barplot_legend(mdf_GP, 
                               cdf_GP, 
                               tax_subgroup)
  my_list <- list(Metadata = pseq_meta,
                  OTUTable = pseq_otu,
                  Taxonomy = pseq_tax,
                  TaxBarplot = main_barplot,
                  TaxLegend = legend)
  return(my_list)
}
```

**Testing my mini-pipeline with neonatal gut microbiome data**
oh my god it worked!!
```{r, fig.width=25,fig.height=15}
## inputs
metadata_FP <- '../data/misc/merged_metadata1.tsv'
seq_depth_FP <- '../data/misc/tss_seq_depth.tsv'
unweighted_FP <- '../data/qiime/core_outputs/unweighted_unifrac_pcoa_results.qza'
weighted_FP <- '../data/qiime/core_outputs/weighted_unifrac_pcoa_results.qza'
faith_pd_FP <- '../data/qiime/core_outputs/faith_pd.tsv'
shannon_FP <- '../data/qiime/core_outputs/shannon_entropy.tsv'
tax_FP <- '../data/qiime/taxonomy.qza'
sample_asv_FP <- '../data/misc/tss_sample_asv.tsv'
otu_table_FP <- '../data/qiime/taxonomy_filtered.qza'
tss_tax_rank <- "Phylum"
tss_tax_subgroup <- "Family"
tss_title <- 'Total Sum Scaled Taxa Barplot'
tss_y <- 'Relative Abundance'
tss_diet_labs <- 
  c('Chow', 
    'High Fat / High Fiber', 
    'High Fat / Low Fiber', 
    'Low Fat / High Fiber', 
    'Low Fat / Low Fiber')

names(tss_diet_labs) <- c('Chow', 
                          'HF/HF', 
                          'HF/LF', 
                          'LF/HF', 
                          'LF/LF')

## barplot function
tss_barplot_files <- microshades_tax_barplot(metadata_FP, 
                                       otu_table_FP, 
                                       tax_FP,
                                       tss_tax_rank, 
                                       tss_tax_subgroup,
                                       tss_title, 
                                       tss_y) 

main_barplot <- tss_barplot_files$TaxBarplot
legend <- tss_barplot_files$TaxLegend

tax_all <- plot_grid(main_barplot, 
                     legend,  
                     rel_widths = c(1, .25),
                     align = 'tblr',
                     axis = 'tblr')
tax_all
```


